---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
library(plyr)
library(dplyr)
library(lattice)

```{r, echo=TRUE}
setwd("C:/Users/Alexander/Documents/Coursera/5. Reproducible Research")

activity <- read.csv("activity.csv")
activity <- tbl_df(activity)
```


## What is mean total number of steps taken per day?
```{r, echo=TRUE}
dayactivity <- group_by(activity, date)

daysum <- summarise(dayactivity, steps = sum(steps, na.rm=T))
hist(daysum$steps, xlab = "Total Steps Per Day", main = "Total Number of Steps Per Day")

print.data.frame(summarise(dayactivity, steps = mean(steps, na.rm=T)))
print.data.frame(summarise(dayactivity, steps = median(steps, na.rm=T)))
```


## What is the average daily activity pattern?
```{r, echo=TRUE}
meanint <- group_by(activity, interval)
avint <- summarise(meanint, steps = mean(steps, na.rm = T))
plot(avint, type = "l")
print(max(avint$steps))
```

## Imputing missing values
```{r, echo=TRUE}
sum(!complete.cases(activity))

mergedint <- merge(meanint, avint, by = "interval")
mergedint$steps.x[is.na(mergedint$steps.x)] <- mergedint$steps.y[is.na(mergedint$steps.x)]
              
q4 <- group_by(mergedint, date)   
q4a <- summarise(q4, steps = sum(steps.x, na.rm=T))
  
hist(q4a$steps, xlab = "Total Steps", main = "Total Number of Steps with Missing Values Filled In" )

print.data.frame(summarise(q4, steps = mean(steps.x, na.rm=T)))
print.data.frame(summarise(q4, steps = median(steps.x, na.rm=T)))
```
By imputing the missing data, we receive a plot which more closely resembles a standard distribution.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
q4$date <- as.POSIXct(q4$date)
daysofweek <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
q4$Weekday <- factor((weekdays(q4$date) %in% daysofweek), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

int2 <- group_by(q4, interval)
avint2 <- summarise(int2, steps = mean(steps.x, na.rm = T))
xyplot(avint2$steps ~ avint2$interval|q4$Weekday, type = "l", layout = c(1,2))
```